Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic

### 2020-09-18 write-up/covid-writeup.tex
Update covid-writeup.tex

Here we consider estimation of the active infection rate in Indiana.  We start with IPW estimation.  Table .

The report does not include

\begin{table}[!th]
\begin{tabular}{c | c | r r | r | r r }
& & \multicolumn{2}{c}{Total Tests (\%)} & US & \multicolumn{2}{c}{Positive Test Rate (\%)}\\
\cline{3-4} \cline{6-7}
& & NonRandom & Random & Census & NonRandom & Random \\ \hline
\multirow{2}{*}{Sex} & Female & 58.2 & 55.0 & 51 & 21.7 & 1.4 \\
& Male & 42 & 45.0 & 49 & 24.2 & 2.1 \\ \hline
\multirow{3}{*}{Age} & $<40$ & 40 & 28.0 & 53 & 29.7 & 1.7 \\
& $40-59$ & 41 & 36.0 & 25 & 24.9 & 2.1 \\
& $\geq 60$ & 19 & 36.0 & 22 & 6.7 & 0.9 \\ \hline
\multirow{2}{*}{Race} & White & 23 & 92.0 & 87 & 19.5 & 1.5 \\
& Nonwhite & 77 & 8.0 & 13 & 25.0 & 3.4 \\ \hline
Hispanic & Hispanic & - & 2.0 & - & 37.6 & 6.9 \\
Origin & Non-Hispanic & - & 98.0 & - & 13.0 & 1.3 \\ \hline
\multirow{2}{*}{Fever} & Yes & 17.0 & 1.8 & - & 66.4 & 4.5 \\
& No & 83.0 & 98.2 & - & 15.6 & 1.3 \\ \hline
\multirow{2}{*}{Household} & Yes & 10.8 & 1.4 & - & 46.1 & 24.4 \\
& No & 89.2 & 98.6 & - & 21.6 & 1.3 \\ \hline
Prior $+$ & Yes & 6.1 & 1.6 & - & 39.2 & 24.4 \\
Test & No & 93.9 & 98.6 & - & 21.6 & 1.3 \\ \hline
\end{tabular}
\caption{Estimated point prevalence of active infection with SARS-CoV-2 by demographics in Indiana.}
\end{table}

>>>>

Here we consider estimation of the active infection rate in Indiana.  Table~\ref{tab:indiana} summarizes data from a random sample from April 25--29th as well as a nonrandom sample obtained between May 2--3 in racial/ethnic minority communities. The nonrandom sample is therefore not a random subsample of the overall case-counts; however, this data provides important supplementary covariate information.  Namely, the samples collected included covariate information on (1) prior symptoms, (2) COVID-19 positive individual in household, and (3) prior COVID-19 positive test. To account for the nonrandom sample being from high risk areas, we adjust estimates to match the statewide positivity rate of 11.7\% between May 1--2. These adjusted rates are presented in parentheses in Table~\ref{tab:indiana}.

\begin{table}[th]
\begin{tabular}{c | c | r r | r | r r }
& & \multicolumn{2}{c}{Total Tests (\%)} & US & \multicolumn{2}{c}{Positive Test Rate (\%)}\\
\cline{3-4} \cline{6-7}
& & NonRandom & Random & Census & NonRandom & Random \\ \hline
\multirow{2}{*}{Sex} & Female & 58.2 & 55.0 & 50.7 & 21.7 (11.2) & 1.4 \\
& Male & 41.8 & 45.0 & 49.3 & 24.2 (12.4) & 2.1 \\ \hline
\multirow{3}{*}{Age} & $<40$ & 39.4 & 28.0 & 52.7 & 29.7 & 1.7 \\
& $40-59$ & 41.1 & 36.0 & 25.2 & 24.9 & 2.1 \\
& $\geq 60$ & 19.5 & 36.0 & 22 & 6.7 & 0.9 \\ \hline
\multirow{2}{*}{Race} & White & 23.1 & 92.0 & 86.9 & 19.5 & 1.5 \\
& Nonwhite & 76.9 & 8.0 & 13.1 & 25.0 & 3.4 \\ \hline
Hispanic & Hispanic & 44.1 & 2.0 & 7.1 & 37.6 & 6.9 \\
Origin & Non-Hispanic & 55.9 & 98.0 & 92.9 & 13.0 & 1.3 \\ \hline
\multirow{2}{*}{Fever} & Yes & 17.0 & 1.8 & - & 66.4 & 4.5 \\
& No & 83.0 & 98.2 & - & 15.6 & 1.3 \\ \hline
Household & Yes & 10.8 & 1.4 & - & 46.1 & 24.4 \\
$+$ Case & No & 89.2 & 98.6 & - & 21.6 & 1.3 \\ \hline
Prior $+$ & Yes & 6.1 & 1.6 & - & 39.2 & 24.4 \\
Test & No & 93.9 & 98.6 & - & 21.6 & 1.3 \\ \hline
\end{tabular}
\caption{Estimated point prevalence of active infection with SARS-CoV-2 by demographics in Indiana.}
\label{tab:indiana}
\end{table}
